---
title: "Installation"
output:
  blogdown::html_page:
    toc: true
weight: 1
---

There are two ways to install dependencies and set programming environment:

* install multiple R packages and dependencies from github, it may take 5-30 mins.

* install [Docker](https://www.docker.com/) first and download [simpipe Docker image](https://hub.docker.com/r/duohongrui/simpipe) from DockerHub. All manipulations in R can be done by [simpipeDoker](https://github.com/duohongrui/simpipe_Docker) package.

## Installation from Github 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "")
```

You can install the development version of simutils, simmethods and simpipe from GitHub with:
```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("duohongrui/simutils")
devtools::install_github("duohongrui/simmethods")
devtools::install_github("duohongrui/simpipe")
```

If you want to use simpipe to simulate datasets with **trajectory information**, please install **tislingshot** first by:
```{r eval=FALSE}
devtools::install_github("dynverse/ti_slingshot/package/")
```

## Download Docker Image

### Install Docker Desktop

If you have not installed official Docker Desktop, please click [here](https://www.docker.com/) and follow the instructions to install it.
{{% notice note %}}
The version of Docker Desktop for Apple Chip is available
{{% /notice %}}

After installing Docker, start Docker service, open an R session and type `dynwrap::test_docker_installation(detailed = TRUE)`:
```{r}
dynwrap::test_docker_installation(detailed = TRUE)
```
If it is `TRUE` in the result, that means Docker is successfully installed.

### Install simpipeDocker


### Download simpipe Docker Image

Use `babelwhale::pull_container("duohongrui/simpipe")` command to pull simpipe Docker image in R (it will take 3-10 mins):
```{r eval=FALSE}
# install.packages("babelwhale")
babelwhale::pull_container("duohongrui/simpipe")
```

Once the image has been downloaded, check the image list:
```{r}
babelwhale::list_docker_images(container_id = "duohongrui/simutils")
```

