---
title: "scDesign"
output:
  blogdown::html_page:
    toc: yes
weight: 11
---

<!-- github markdown built using 
rmarkdown::render("vignettes/scDesign.Rmd", output_format = rmarkdown::github_document())
-->

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#", collapse = TRUE)
```
Here scDesign method will be demonstrated clearly and hope that this document can help you.


## Simulating datasets using scDesign

There is no estimation step when using scDesign, so we can directly simulate new datasets through reference data.

```{r load_data, message=FALSE, warning=FALSE, error=FALSE}
library(simmethods)
library(SingleCellExperiment)
# Load data
ref_data <- simmethods::data
```

We will simulate a dataset based on refernece data with different scenarios.

1. Datasets with default parameters
2. Determin the number of cells and genes
3. Simulate two or more groups

### Datasets with default parameters

The reference data contains 160 cells and 4000 genes, if we simulate datasets with default parameters and then we will obtain a new data which has the same size as the reference data. In addtion, the simulated dataset will have one group of cells.

```{r default_params}
simulate_result <- simmethods::scDesign_simulation(
  ref_data = ref_data,
  return_format = "SCE",
  seed = 111
)
```

```{r}
SCE_result <- simulate_result[["simulate_result"]]
dim(SCE_result)
```

```{r}
head(colData(SCE_result))
```


### Determin the number of cells

{{% notice note %}}
We can only set the cell number in scDesign.
{{% /notice %}}

Here, we simulate a new dataset with 500 cells:
```{r customed cell and gene number}
simulate_result <- simmethods::scDesign_simulation(
  ref_data = ref_data,
  return_format = "list",
  other_prior = list(nCells = 500),
  seed = 111
)
```

```{r}
result <- simulate_result[["simulate_result"]][["count_data"]]
dim(result)
```


### Simulate two or more groups

In scDesign, we can et `nGroups` directly, together with the proportions of different cell groups by `prob.group`. Moreover, the proportion of DEGs via `de.prob` and fold change via `fc.group` can be customed.

For demonstration, we will simulate three groups.
```{r groups}
simulate_result <- simmethods::scDesign_simulation(
  ref_data = ref_data,
  return_format = "list",
  other_prior = list(nCells = 500,
                     nGroups = 3,
                     prob.group = c(0.1, 0.3, 0.6),
                     de.prob = 0.2,
                     fc.group = 4),
  seed = 111
)
```

```{r}
result <- simulate_result[["simulate_result"]][["count_data"]]
dim(result)
```


```{r}
## cell information
cell_info <- simulate_result[["simulate_result"]][["col_meta"]]
table(cell_info$group)
```

```{r}
## gene information
gene_info <- simulate_result[["simulate_result"]][["row_meta"]]
### the proportion of DEGs
table(gene_info$de_gene)[2]/nrow(result) ## de.prob = 0.2
```
