---
title: "BEARscc"
output:
  blogdown::html_page:
    toc: yes
weight: 18
---

<!-- github markdown built using 
rmarkdown::render("vignettes/BEARscc.Rmd", output_format = rmarkdown::github_document())
-->

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#", collapse = TRUE)
```
Here BEARscc method will be demonstrated clearly and hope that this document can help you.

## Estimating parameters from a real dataset

Before simulating datasets, it is important to estimate some essential parameters from a real dataset in order to make the simulated data more real. The reference data can be downloaded [here](https://zenodo.org/record/8251596/files/data23_GSE62270.rds?download=1).

BEARscc needs spike-in ERCC genes as the reference to measure the variation of the
real dataset and there are some notes that users must pay attention to.

1. Make sure that there are spike-in genes in your count matrix whose prefix are `ERCC-`.
If not, the error may occur.

2. BEARscc needs ensembl gene id to execute estimation step, so it is better
to transform the gene id previously. But users can also input official gene id
and the procedure will convert them into ensembl gene id and note that this
step may result in losing some genes when matching gene ids.

3. If users need the transformation of gene ids, users must input the `species`
name parameter: `mouse` or `human`. And we will match the according database to
accomplish the conversion step.

4. Another important parameters: `dilution.factor`, `volume`
- `dilution.factor`: The dilution factor to dilute the ERCC spike-in mix liquid.
- `volume`: The volume (microliter) of spike-in mix used in sequencing step.

```{r load_data, message=FALSE, warning=FALSE, error=FALSE}
library(simmethods)
library(SingleCellExperiment)
# Load data (downloaded from https://zenodo.org/record/8251596/files/data23_GSE62270.rds?download=1)
data <- readRDS("../../../../preprocessed_data/data23_GSE62270.rds")
ref_data <- data$data
## group_condition can must be a numeric vector.
other_prior <- list(dilution.factor = 50000, volume = 0.03, species = "mouse")
```

Using `simmethods::BEARscc_estimation` command to execute the estimation step.

```{r estimate, warning=FALSE}
estimate_result <- simmethods::BEARscc_estimation(
  ref_data = ref_data,
  other_prior = other_prior,
  verbose = TRUE,
  seed = 8
)
```

## Simulating datasets using BEARscc

{{% notice note %}}
Users can not set the number of cells or genes in BEARscc.
{{% /notice %}}

```{r default_params}
simulate_result <- simmethods::BEARscc_simulation(
  parameters = estimate_result[["estimate_result"]],
  return_format = "list",
  seed = 111
)
```


```{r}
result <- simulate_result[["simulate_result"]]
dim(result$count_data)
```
