---
title: "SPsimSeq"
output:
  blogdown::html_page:
    toc: yes
weight: 19
---

<!-- github markdown built using 
rmarkdown::render("vignettes/SPsimSeq.Rmd", output_format = rmarkdown::github_document())
-->

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#", collapse = TRUE)
```

There is no individual estimation step using SPsimSeq as the estimation is combined with simulation step.

```{r load_data, message=FALSE, warning=FALSE, error=FALSE}
library(simmethods)
library(SingleCellExperiment)
# Load data
ref_data <- simmethods::data
```


## Simulating datasets using SPsimSeq

1. Datasets with default parameters
2. Determin the number of cells and genes
3. Simulate groups
4. Simulate batches

### Datasets with default parameters

The reference data contains 160 cells and 4000 genes, if we simulate datasets with default parameters and then we will obtain a new data which has the same size as the reference data.


```{r default_params}
simulate_result <- simmethods::SPsimSeq_simulation(
  ref_data = ref_data,
  other_prior = NULL,
  return_format = "SCE",
  seed = 111
)
```

```{r}
SCE_result <- simulate_result[["simulate_result"]]
dim(SCE_result)
```

```{r}
head(colData(SCE_result))
```

```{r}
head(rowData(SCE_result))
```

### Determin the number of cells and genes

```{r}
simulate_result <- simmethods::SPsimSeq_simulation(
  ref_data = ref_data,
  other_prior = list(nCells = 500,
                     nGenes = 1000),
  return_format = "SCE",
  seed = 111
)
```

```{r}
SCE_result <- simulate_result[["simulate_result"]]
dim(SCE_result)
```


### Simulate groups

The number of groups simulated by SPsimSeq is determined by the group information used in simulation step. If the group information of two cell states is provided, then the simulated dataset will contain two groups. SPsimSeq also provides other parameters related to DEGs such as the proportion of DEGs (`de.prob`) and the fold change of DGEs (`fc.group`).


For demonstration, we will simulate two groups using the learned parameters. We can set `de.prob = 0.2` to simulate 20% genes as DEGs.
```{r}
group_condition <- as.numeric(simmethods::group_condition)
simulate_result <- simmethods::SPsimSeq_simulation(
  ref_data = ref_data,
  other_prior = list(group.condition = group_condition,
                     de.prob = 0.2,
                     fc.group = 4),
  return_format = "list",
  seed = 111
)
```


```{r}
col_data <- simulate_result[["simulate_result"]][["col_meta"]]
table(col_data$group)
```

```{r}
## gene information
gene_info <- simulate_result[["simulate_result"]][["row_meta"]]
### the proportion of DEGs
table(gene_info$de_gene)[2]/nrow(gene_info) ## de.prob = 0.2
```


### Simulate batches

Users can simulate batches when `batch.condition` parameter is activated and just input the numeric vectors that specify the batch labels of cells.

```{r}
set.seed(111)
ref_data <- scater::mockSCE(ncells = 160,
                            ngenes = 4000)
ref_data <- SingleCellExperiment::counts(ref_data)
```


For demonstration, we will simulate two batches.
```{r}
simulate_result <- simmethods::SPsimSeq_simulation(
  ref_data = ref_data,
  other_prior = list(batch.condition = sample(1:3, 160, replace = TRUE)),
  return_format = "list",
  seed = 111
)
```


```{r}
## cell information
col_data <- simulate_result[["simulate_result"]][["col_meta"]]
```

```{r}
table(col_data$batch)
```
