---
title: "muscat"
output:
  blogdown::html_page:
    toc: yes
weight: 22
---

<!-- github markdown built using 
rmarkdown::render("vignettes/muscat.Rmd", output_format = rmarkdown::github_document())
-->

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#", collapse = TRUE)
```
Here muscat method will be demonstrated clearly and hope that this document can help you.

## Estimating parameters from a real dataset

Before simulating datasets, it is important to estimate some essential parameters from a real dataset in order to make the simulated data more real. 

```{r load_data, message=FALSE, warning=FALSE, error=FALSE}
library(simmethods)
library(SingleCellExperiment)
# Load data
ref_data <- simmethods::data
```

### Default eatimation

```{r, warning=FALSE}
estimate_result <- simmethods::muscat_estimation(
  ref_data = ref_data,
  other_prior = NULL,
  verbose = T,
  seed = 10
)
```

### Estimation with cell groups

When you use muscat to estimate parameters from a real dataset, you can also input a numeric vector to specify the groups or plates that each cell comes from, like `other_prior = list(group.condition = the numeric vector)`.

```{r estimate, warning=FALSE}
group_condition <- as.numeric(simmethods::group_condition)
estimate_result <- simmethods::muscat_estimation(
  ref_data = ref_data,
  other_prior = list(group.condition = group_condition),
  verbose = T,
  seed = 10
)
```

## Simulating datasets using muscat

After estimating parameter from a real dataset, we will simulate a dataset based on the learned parameters with different scenarios.

1. Datasets with default parameters
2. Determin the number of cells and genes
3. Simulate two or more groups

### Datasets with default parameters

The reference data contains 160 cells and 4000 genes, if we simulate datasets with default parameters and then we will obtain a new data which has the same size as the reference data. In addtion, the simulated dataset will have one group of cells.

```{r default_params}
simulate_result <- simmethods::muscat_simulation(
  parameters = estimate_result[["estimate_result"]],
  other_prior = NULL,
  return_format = "SCE",
  seed = 111
)
```

```{r}
SCE_result <- simulate_result[["simulate_result"]]
dim(SCE_result)
```

### Determin the number of cells and genes

In muscat, we can set `nCells` and `nGenes` to specify the number of cells and genes.

Here, we simulate a new dataset with 1000 cells and 1000 genes:

```{r customed cell and gene number}
simulate_result <- simmethods::muscat_simulation(
  parameters = estimate_result[["estimate_result"]],
  return_format = "list",
  other_prior = list(nCells = 1000,
                     nGenes = 1000),
  seed = 111
)
```

```{r}
result <- simulate_result[["simulate_result"]][["count_data"]]
dim(result)
```

### Simulate two or more groups

In muscat, we can set `nGroups` directly to specify the number of simulated groups. muscat also provides other parameters related to DEGs such as the proportion of DEGs (`de.prob`) and the fold change of DGEs (`fc.group`).

For demonstration, we will simulate two groups using the learned parameters.
```{r groups}
simulate_result <- simmethods::muscat_simulation(
  parameters = estimate_result[["estimate_result"]],
  return_format = "list",
  other_prior = list(nCells = 500,
                     nGenes = 1000,
                     nGroups = 2,
                     de.prob = 0.4,
                     fc.group = 4),
  seed = 111
)
```

```{r}
result <- simulate_result[["simulate_result"]][["count_data"]]
dim(result)
```

```{r}
## cell information
cell_info <- simulate_result[["simulate_result"]][["col_meta"]]
table(cell_info$group)
```

```{r}
## gene information
gene_info <- simulate_result[["simulate_result"]][["row_meta"]]
### the proportion of DEGs
table(gene_info$de_gene)[2]/nrow(result) ## de.prob = 0.4
```
