---
title: "PROSSTT"
output:
  blogdown::html_page:
    toc: yes
weight: 26
---

<!-- github markdown built using 
rmarkdown::render("vignettes/PROSSTT.Rmd", output_format = rmarkdown::github_document())
-->

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#", collapse = TRUE)
```
Here PROSSTT method will be demonstrated clearly and hope that this document can help you.

## Estimating parameters from a real dataset

Before simulating datasets, it is important to estimate some essential parameters from a real dataset in order to make the simulated data more real. 

```{r load_data, message=FALSE, warning=FALSE, error=FALSE}
library(simmethods)
library(SingleCellExperiment)
# Load data
ref_data <- simmethods::data
```


```{r, warning=FALSE}
estimate_result <- simmethods::PROSSTT_estimation(
  ref_data = ref_data,
  verbose = T,
  seed = 10
)
```

See the result:
```{r}
estimate_result[["estimate_result"]][["newick_tree"]]
```
You can obtain a tree structure of Newick format where cells can be sampled from to generate the datasets with trajectory. Then the hierarchical clustering is used to obtain the relationship between different groups. If no group information is provided like above codes, the groups or clusters are determined by Seurat pipeline.

Users can also input the group information of cells:

```{r}
group <- as.numeric(simmethods::group_condition)
estimate_result <- simmethods::PROSSTT_estimation(
  ref_data = ref_data,
  other_prior = list(group.condition = group),
  verbose = T,
  seed = 10
)
```


## Simulating datasets with cell trajectory using PROSSTT

After estimating parameter from a real dataset, we will simulate a dataset based on the learned parameters with different scenarios.

1. Datasets with default parameters
2. Determin the number of cells and genes
3. Visualization

### Datasets with default parameters

The reference data contains 160 cells and 4000 genes, if we simulate datasets with default parameters and then we will obtain a new data which has the same size as the reference data.

```{r default_params}
simulate_result <- simmethods::PROSSTT_simulation(
  parameters = estimate_result[["estimate_result"]],
  other_prior = NULL,
  return_format = "SCE",
  seed = 111
)
```

```{r}
SCE_result <- simulate_result[["simulate_result"]]
dim(SCE_result)
```

### Determin the number of cells and genes

In PROSSTT, we can set `nCells` and `nGenes` to specify the number of cells and genes.

Here, we simulate a new dataset with 1000 cells and 1000 genes:

```{r customed cell and gene number}
simulate_result <- simmethods::PROSSTT_simulation(
  parameters = estimate_result[["estimate_result"]],
  return_format = "list",
  other_prior = list(nCells = 1000,
                     nGenes = 1000),
  seed = 111
)
```

```{r}
result <- simulate_result[["simulate_result"]][["count_data"]]
dim(result)
```


### Visualization

Make sure that you have already installed several R packages:
```{r}
if(!requireNamespace("dynwrap", quietly = TRUE)){install.packages("dynwrap")}
if(!requireNamespace("dyndimred", quietly = TRUE)){install.packages("dyndimred")}
if(!requireNamespace("dynplot", quietly = TRUE)){install.packages("dynplot")}
if(!requireNamespace("tislingshot", quietly = TRUE)){devtools::install_github("dynverse/ti_slingshot/package/")}
```

First we should wrap the data into a standard object:

```{r}
dyn_object <- dynwrap::wrap_expression(counts = t(result),
                                       expression = log2(t(result) + 1))
```

Next, we infer the trajectory using SlingShot which has been proved to be the most best method to do this:

```{r}
model <- dynwrap::infer_trajectory(dataset = dyn_object,
                                   method = tislingshot::ti_slingshot(),
                                   parameters = NULL,
                                   give_priors = NULL,
                                   seed = 111,
                                   verbose = TRUE)
```

Finally, we can plot the trajectory after performing dimensionality reduction:
```{r, fig.align='center', fig.height=6, fig.width=12}
dimred <- dyndimred::dimred_umap(dyn_object$expression)
dynplot::plot_dimred(model, dimred = dimred)
```

For more details about trajectory inference and visualization, please check [dynverse](https://dynverse.org/).