---
title: "dyntoy"
output:
  blogdown::html_page:
    toc: yes
weight: 28
---

<!-- github markdown built using 
rmarkdown::render("vignettes/dyntoy.Rmd", output_format = rmarkdown::github_document())
-->

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#", collapse = TRUE)
```
Here dyntoy method will be demonstrated clearly and hope that this document can help you.

## Estimating parameters from a real dataset

Before simulating datasets, it is important to estimate some essential parameters from a real dataset in order to make the simulated data more real. 

```{r load_data, message=FALSE, warning=FALSE, error=FALSE}
library(simmethods)
library(SingleCellExperiment)
# Load data
ref_data <- simmethods::data
```

To simulate trajectory datasets using dyntoy, we first performe trajectory inference analysis to the real data. If no group information is provided, the k-means algorithm is used to determine the number of clusters(groups) that the real data contains.

```{r, warning=FALSE}
estimate_result <- simmethods::dyntoy_estimation(
  ref_data = ref_data,
  verbose = T,
  seed = 10
)
```

Users can also input the group information of cells and the k-means will not be used:

```{r}
group <- as.numeric(simmethods::group_condition)
estimate_result <- simmethods::dyntoy_estimation(
  ref_data = ref_data,
  other_prior = list(group.condition = group),
  verbose = T,
  seed = 10
)
```


## Simulating datasets with cell trajectory using dyntoy

After estimating parameter from a real dataset, we will simulate a dataset based on the learned parameters with different scenarios.

1. Datasets with default parameters
2. Determin the number of cells and genes
3. Visualization

### Datasets with default parameters

The reference data contains 160 cells and 4000 genes, if we simulate datasets with default parameters and then we will obtain a new data which has the same size as the reference data.

```{r default_params}
simulate_result <- simmethods::dyntoy_simulation(
  parameters = estimate_result[["estimate_result"]],
  other_prior = NULL,
  return_format = "SCE",
  seed = 111
)
```

```{r}
SCE_result <- simulate_result[["simulate_result"]]
dim(SCE_result)
```

### Determin the number of cells and genes

In dyntoy, we can set `nCells` and `nGenes` to specify the number of cells and genes.

Here, we simulate a new dataset with 1000 cells and 1000 genes:

```{r customed cell and gene number, warning=FALSE}
simulate_result <- simmethods::dyntoy_simulation(
  parameters = estimate_result[["estimate_result"]],
  return_format = "list",
  other_prior = list(nCells = 1000,
                     nGenes = 1000),
  seed = 111
)
```

```{r}
result <- simulate_result[["simulate_result"]][["count_data"]]
dim(result)
```


### Visualization

Make sure that you have already installed several R packages:
```{r}
if(!requireNamespace("dynwrap", quietly = TRUE)){install.packages("dynwrap")}
if(!requireNamespace("dyndimred", quietly = TRUE)){install.packages("dyndimred")}
if(!requireNamespace("dynplot", quietly = TRUE)){install.packages("dynplot")}
if(!requireNamespace("tislingshot", quietly = TRUE)){devtools::install_github("dynverse/ti_slingshot/package/")}
```

First we should wrap the data into a standard object:

```{r}
dyn_object <- dynwrap::wrap_expression(counts = t(result),
                                       expression = log2(t(result) + 1))
```

Next, we infer the trajectory using SlingShot which has been proved to be the most best method to do this:

```{r}
model <- dynwrap::infer_trajectory(dataset = dyn_object,
                                   method = tislingshot::ti_slingshot(),
                                   parameters = NULL,
                                   give_priors = NULL,
                                   seed = 111,
                                   verbose = TRUE)
```

Finally, we can plot the trajectory after performing dimensionality reduction:
```{r, fig.align='center', fig.height=6, fig.width=12}
dimred <- dyndimred::dimred_umap(dyn_object$expression)
dynplot::plot_dimred(model, dimred = dimred)
```

For more details about trajectory inference and visualization, please check [dynverse](https://dynverse.org/).