---
title: "scMultiSim"
output:
  blogdown::html_page:
    toc: yes
weight: 31
---

<!-- github markdown built using 
rmarkdown::render("vignettes/scMultiSim.Rmd", output_format = rmarkdown::github_document())
-->

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#", collapse = TRUE)
```

Here scMultiSim method will be demonstrated clearly and hope that this document can help you.

## Estimating parameters from a real dataset

Before simulating datasets, it is important to estimate some essential parameters from a real dataset in order to make the simulated data more real. 

```{r load_data, message=FALSE, warning=FALSE, error=FALSE}
library(simmethods)
library(SingleCellExperiment)
# Load data
ref_data <- simmethods::data
```


```{r, warning=FALSE}
estimate_result <- simmethods::scMultiSim_estimation(
  ref_data = ref_data,
  verbose = T,
  seed = 10
)
```

Users can also input the group information of cells:

```{r}
group <- as.numeric(simmethods::group_condition)
estimate_result <- simmethods::scMultiSim_estimation(
  ref_data = ref_data,
  other_prior = list(group.condition = group),
  verbose = T,
  seed = 10
)
```

The estimation result contains an object with `phylo` data structure:

```{r fig.align='center', fig.height=4, fig.width=5}
plot(estimate_result[["estimate_result"]][["phylo"]])
```


## Simulating datasets using scMultiSim

1. Datasets with default parameters
2. Determin the number of cells and genes
3. Simulate groups
4. Simulate batches
5. Simulate cellular differential trajectory

### Datasets with default parameters

The reference data contains 160 cells and 4000 genes, if we simulate datasets with default parameters and then we will obtain a new data which has the same size as the reference data.


```{r default_params}
simulate_result <- simmethods::scMultiSim_simulation(
  parameters = estimate_result$estimate_result,
  other_prior = NULL,
  return_format = "SCE",
  seed = 111
)
```


```{r}
SCE_result <- simulate_result[["simulate_result"]]
dim(SCE_result)
```

```{r}
head(colData(SCE_result))
```

### Determin the number of cells and genes

```{r}
simulate_result <- simmethods::scMultiSim_simulation(
  parameters = estimate_result$estimate_result,
  other_prior = list(nCells = 500,
                     nGenes = 1000),
  return_format = "SCE",
  seed = 111
)
```

```{r}
SCE_result <- simulate_result[["simulate_result"]]
dim(SCE_result)
```


### Simulate groups

In scMultiSim, we can not set `nGroups` directly and should set `prob.group` instead. For example, if we want to simulate 2 groups, we can type `other_prior = list(prob.group = c(0.5, 0.5))`. Note that the sum of `prob.group` numeric vector must equal to 1, so we can also set `prob.group = c(0.3, 0.7)`.

{{% notice note %}}
The group number should be equal to that used or detected in the estimation step, otherwise the error may occur.
{{% /notice %}}


In addtion, if we want to simulate three or more groups, we should obey the rules:

* The length of `prob.group` vector must always equal to the number of groups.
* The sum of `prob.group` numeric vector must equal to 1.

For demonstration, we will simulate two groups using the learned parameters.
```{r}
simulate_result <- simmethods::scMultiSim_simulation(
  parameters = estimate_result$estimate_result,
  other_prior = list(prob.group = c(0.4, 0.6)),
  return_format = "list",
  seed = 111
)
```


```{r}
col_data <- simulate_result[["simulate_result"]][["col_meta"]]
table(col_data$group)
```


### Simulate batches

In scMultiSim, we can set `nBatches` directly to simulate cell batches.

For demonstration, we will simulate two batches.
```{r}
simulate_result <- simmethods::scMultiSim_simulation(
  parameters = estimate_result$estimate_result,
  other_prior = list(nBatches = 2),
  return_format = "list",
  seed = 111
)
```


```{r}
## cell information
col_data <- simulate_result[["simulate_result"]][["col_meta"]]
```


```{r}
table(col_data$batch)
```


### Simulate cellular differential trajectory

The parameter estimation step is the same as that demonstrated above. If you want to simulate datasets with trajectory, you should specify the parameter `traj` in `other_prior` to `TRUE`.


```{r scMultiSim-traj}
simulate_result <- simmethods::scMultiSim_simulation(
  parameters = estimate_result$estimate_result,
  other_prior = list(traj = TRUE),
  return_format = "list",
  seed = 111
)
```


Before visualization, Make sure that you have already installed several R packages:
```{r}
if(!requireNamespace("dynwrap", quietly = TRUE)){install.packages("dynwrap")}
if(!requireNamespace("dyndimred", quietly = TRUE)){install.packages("dyndimred")}
if(!requireNamespace("dynplot", quietly = TRUE)){install.packages("dynplot")}
if(!requireNamespace("tislingshot", quietly = TRUE)){devtools::install_github("dynverse/ti_slingshot/package/")}
```

First we should wrap the data into a standard object:

```{r}
data <- simulate_result$simulate_result$count_data
dyn_object <- dynwrap::wrap_expression(counts = t(data),
                                       expression = log2(t(data) + 1))
```

Next, we infer the trajectory using SlingShot which has been proved to be the most best method to do this:

```{r}
model <- dynwrap::infer_trajectory(dataset = dyn_object,
                                   method = tislingshot::ti_slingshot(),
                                   parameters = NULL,
                                   give_priors = NULL,
                                   seed = 111,
                                   verbose = TRUE)
```

Finally, we can plot the trajectory after performing dimensionality reduction:
```{r, fig.align='center', fig.height=6, fig.width=10}
dimred <- dyndimred::dimred_umap(dyn_object$expression)
dynplot::plot_dimred(model, dimred = dimred)
```

For more details about trajectory inference and visualization, please check [dynverse](https://dynverse.org/).