---
title: "dyngen"
output:
  blogdown::html_page:
    toc: yes
weight: 33
---

<!-- github markdown built using 
rmarkdown::render("vignettes/dyngen.Rmd", output_format = rmarkdown::github_document())
-->

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#", collapse = TRUE)
```
Here dyngen method will be demonstrated clearly and hope that this document can help you.

## Estimating parameters from a real dataset

Before simulating datasets, it is important to estimate some essential parameters from a real dataset in order to make the simulated data more real. 
```{r load_data, message=FALSE, warning=FALSE, error=FALSE}
library(simmethods)
# Load data (downloaded from https://zenodo.org/record/8251596/files/data82_cellbench-SC1_luyitian.rds?download=1)
data <- readRDS("../../../../preprocessed_data/data82_cellbench-SC1_luyitian.rds")
ref_data <- t(as.matrix(data$data$counts))
```

### Default estimation

```{r, warning=FALSE}
estimate_result <- simmethods::dyngen_estimation(
  ref_data = ref_data,
  other_prior = NULL,
  verbose = TRUE,
  seed = 111
)
```


### Information of cell groups

If the information of cell groups is available, you can use another way to estimate the parameters.

```{r}
## cell groups
group_condition <- as.numeric(data$data_info$group_condition)
```


```{r estimate, warning=FALSE}
estimate_result <- simmethods::dyngen_estimation(
  ref_data = ref_data,
  other_prior = list(group.condition = group_condition),
  verbose = TRUE,
  seed = 111
)
```


## Simulating datasets using dyngen

After estimating parameter from a real dataset, we will simulate a dataset based on the learned parameters with different scenarios.

1. Datasets with default parameters
2. Determin the number of cells and genes

### Datasets with default parameters

The reference data contains 157 cells and 1770 genes, if we simulate datasets with default parameters and then we will obtain a new data which has the same size as the reference data.

```{r default_params, warning=FALSE}
simulate_result <- simmethods::dyngen_simulation(
  parameters = estimate_result[["estimate_result"]],
  return_format = "SCE",
  seed = 111
)
```

```{r}
SCE_result <- simulate_result[["simulate_result"]]
dim(SCE_result)
```

### Determin the number of cells and genes

In dyngen, we can set `nCells` and `nGenes` parameters to specify the number of cells and genes that need to be simulated.
Here, we simulate a new dataset with 100 cells and 100 genes:
```{r customed cell and gene number, warning=FALSE}
simulate_result <- simmethods::dyngen_simulation(
  parameters = estimate_result[["estimate_result"]],
  return_format = "list",
  other_prior = list(nCells = 100,
                     nGenes = 100),
  seed = 111
)
```

```{r}
result <- simulate_result[["simulate_result"]][["count_data"]]
dim(result)
```

{{% notice note %}}
Dyngen may need a large amount of memory when simulating new datasets, so users should always focus on your occupied computational resources.
{{% /notice %}}