---
title: "phenopath"
output:
  blogdown::html_page:
    toc: yes
weight: 33
---

<!-- github markdown built using 
rmarkdown::render("vignettes/phenopath.Rmd", output_format = rmarkdown::github_document())
-->

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#", collapse = TRUE)
```
Here phenopath method will be demonstrated clearly and hope that this document can help you.

## Estimating parameters from a real dataset

Before simulating datasets, it is important to estimate some essential parameters from a real dataset in order to make the simulated data more real. 

```{r load_data, message=FALSE, warning=FALSE, error=FALSE}
library(simmethods)
library(SingleCellExperiment)
# Load data
ref_data <- simmethods::data
```

MAF can only simulate datasets with bifurcation trajectory, so the estimation step may fail due to the intrinsic characteristics of real data.

```{r, warning=FALSE}
estimate_result <- simmethods::phenopath_estimation(
  ref_data = ref_data,
  verbose = T,
  seed = 10
)
```


## Simulating datasets with cell trajectory using phenopath

After estimating parameter from a real dataset, we will simulate a dataset based on the learned parameters with different scenarios.

1. Datasets with default parameters
2. Determin the number of cells and genes
3. Visualization

### Datasets with default parameters

The reference data contains 160 cells and 4000 genes, if we simulate datasets with default parameters and then we will obtain a new data which has the same size as the reference data.

```{r default_params, warning=FALSE}
simulate_result <- simmethods::phenopath_simulation(
  parameters = estimate_result[["estimate_result"]],
  other_prior = NULL,
  return_format = "SCE",
  seed = 111
)
```

```{r}
SCE_result <- simulate_result[["simulate_result"]]
dim(SCE_result)
```

### Determin the number of cells and genes

In phenopath, we can set `nCells` and `nGenes` to specify the number of cells and genes.

Here, we simulate a new dataset with 2000 cells and 2000 genes:

```{r customed cell and gene number, warning=FALSE}
simulate_result <- simmethods::phenopath_simulation(
  parameters = estimate_result[["estimate_result"]],
  return_format = "list",
  other_prior = list(nCells = 2000,
                     nGenes = 2000),
  seed = 11
)
```

```{r}
result <- simulate_result[["simulate_result"]][["count_data"]]
dim(result)
```


### Visualization

Make sure that you have already installed several R packages:
```{r}
if(!requireNamespace("dynwrap", quietly = TRUE)){install.packages("dynwrap")}
if(!requireNamespace("dyndimred", quietly = TRUE)){install.packages("dyndimred")}
if(!requireNamespace("dynplot", quietly = TRUE)){install.packages("dynplot")}
if(!requireNamespace("tislingshot", quietly = TRUE)){devtools::install_github("dynverse/ti_slingshot/package/")}
```

First we should wrap the data into a standard object:

```{r}
dyn_object <- dynwrap::wrap_expression(counts = t(result),
                                       expression = log2(t(result) + 1))
```

Next, we infer the trajectory using SlingShot which has been proved to be the most best method to do this:

```{r, eval=FALSE}
model <- dynwrap::infer_trajectory(dataset = dyn_object,
                                   method = tislingshot::ti_slingshot(),
                                   parameters = NULL,
                                   give_priors = NULL,
                                   seed = 111,
                                   verbose = TRUE)
```

However, the trajectory inference failed because of the inference method itself and we can choose MST method to try again.

Open your Docker Desktop and check your Docker execution status. If you do not install Docker, please turn to [Docker homepage](https://www.docker.com/).

```{r}
dynwrap::test_docker_installation(detailed = TRUE)
```


```{r, warning=FALSE}
model <- dynwrap::infer_trajectory(dataset = dyn_object,
                                   method = "mst",
                                   parameters = NULL,
                                   give_priors = NULL,
                                   seed = 111,
                                   verbose = FALSE)
```

Finally, we can plot the trajectory after performing dimensionality reduction:
```{r, fig.align='center', fig.height=6, fig.width=12}
dimred <- dyndimred::dimred_umap(dyn_object$expression)
dynplot::plot_dimred(model, dimred = dimred)
```

For more details about trajectory inference and visualization, please check [dynverse](https://dynverse.org/).