---
title: "Lun2"
output:
  blogdown::html_page:
    toc: yes
weight: 6
---

<!-- github markdown built using 
rmarkdown::render("vignettes/Lun2.Rmd", output_format = rmarkdown::github_document())
-->

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "#", collapse = TRUE)
```
Here Lun2 method will be demonstrated clearly and hope that this document can help you.

## Estimating parameters from a real dataset

Before simulating datasets, it is important to estimate some essential parameters from a real dataset in order to make the simulated data more real. If you do not have a single-cell transcriptomics count matrix now, you can use the data collected in simmethods package by `simmethods:data` command.

When you use Lun2 to estimate parameters from a real dataset, you must input a numeric vector to specify the batches or plates that each cell comes from, like `other_prior = list(batch.condition = the numeric vector)`.

```{r load_data, message=FALSE, warning=FALSE, error=FALSE}
library(simmethods)
library(SingleCellExperiment)
# Load data
ref_data <- simmethods::data
plates <- simmethods::group_condition
## plates can must be a numeric vector.
other_prior <- list(batch.condition = as.numeric(plates))
```

Using `simmethods::Lun2_estimation` command to execute the estimation step.

```{r estimate}
estimate_result <- simmethods::Lun2_estimation(ref_data = ref_data,
                                               other_prior = other_prior,
                                               verbose = T,
                                               seed = 10)
```

## Simulating datasets using Lun2

After estimating parameter from a real dataset, we will simulate a dataset based on the learned parameters with different scenarios.

1. Datasets with default parameters
2. Determin the number of cells and genes
3. Simulate two or more batches

### Datasets with default parameters

The reference data contains 160 cells and 4000 genes, if we simulate datasets with default parameters and then we will obtain a new data which has the same size as the reference data. In addtion, the simulated dataset will have one batch of cells.

```{r default_params}
simulate_result <- simmethods::Lun2_simulation(
  parameters = estimate_result[["estimate_result"]],
  return_format = "SCE",
  seed = 111
)
```

```{r}
SCE_result <- simulate_result[["simulate_result"]]
dim(SCE_result)
```

```{r}
head(colData(SCE_result))
```

```{r}
head(rowData(SCE_result))
```


### Determin the number of cells and genes

In Lun2, we can not set `nCells` directly and should set `cell.plates` instead. For example, if we want to simulate 500 cells, we can type `other_prior = list(cell.plates = rep(1, 500))`. For genes, we can just set `nGenes`.
Here, we simulate a new dataset with 500 cells and 2000 genes:
```{r customed cell and gene number}
simulate_result <- simmethods::Lun2_simulation(
  parameters = estimate_result[["estimate_result"]],
  return_format = "list",
  other_prior = list(cell.plates = rep(1, 500),
                     nGenes = 2000),
  seed = 111
)
```

{{% notice note %}}
The cell.plates parameter represents the sampling source of cells in real experiments.
{{% /notice %}}

```{r}
result <- simulate_result[["simulate_result"]][["count_data"]]
dim(result)
```


### Simulate two or more batches

In Lun2, we can not set `nbatches` directly and should set `cell.plates ` instead. For example, if we want to simulate 2 batches, we can type `other_prior = list(cell.plates  = sample(1:2, 500, replace = TRUE))`. Note that the length of `cell.plates ` numeric vector must be equal to the cell number.

For demonstration, we will simulate three batches using the learned parameters.
```{r batches}
simulate_result <- simmethods::Lun2_simulation(
  parameters = estimate_result[["estimate_result"]],
  return_format = "list",
  other_prior = list(cell.plates = sample(1:2, 500, replace = TRUE),
                     nGenes = 2000),
  seed = 111
)
```

```{r}
result <- simulate_result[["simulate_result"]][["count_data"]]
dim(result)
```


```{r}
## cell information
cell_info <- simulate_result[["simulate_result"]][["col_meta"]]
table(cell_info$plate)
```
