---
title: "Installation"
output:
  blogdown::html_page:
    toc: true
weight: 1
---

There are two ways to configure programming environment and execute simulation procedure:

* install multiple R packages and dependencies from github, it may take 5-30 mins.

* install [Docker](https://www.docker.com/) first and download [simpipe Docker image](https://hub.docker.com/repository/docker/duohongrui/simpipe/general) from DockerHub. All manipulations in R can be done by a lightweight [simpipe2doker](https://github.com/duohongrui/simpipe2docker) package.

## Choice1: Installation from Github 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE, comment = "")
```
The first way to configure your programming environment is to install all of the packages and necessary dependencies 

You can install the development version of **simutils**, **simmethods** and **simpipe** from GitHub by:
```{r eval=FALSE}
# install.packages("devtools")
devtools::install_github("duohongrui/simutils")
devtools::install_github("duohongrui/simmethods")
devtools::install_github("duohongrui/simpipe")
```

If you want to use simpipe to simulate datasets with **trajectory information**, please install **tislingshot** first by:
```{r eval=FALSE}
devtools::install_github("dynverse/ti_slingshot/package/")
```


## Choice2: Download Docker Image
We provide another way for users to configure your environment by Docker. To achieve the interaction between Docker and R environment, we build a lightweight R package [simpipe2docker](https://github.com/duohongrui/simpipe2docker) which is used for calling commands in terminal to link Docker and R. It is more easier to manipulate and suitable for different operation systems and R versions, avoiding complex configurations in various environments.

### Install Docker Desktop

If you have not installed official Docker Desktop, please click [here](https://www.docker.com/) and follow the [instructions](https://docs.docker.com/desktop/install/mac-install/) to install it.
{{% notice note %}}
The version of Docker Desktop for Apple Chip is available
{{% /notice %}}

After installing Docker, start Docker service, open an R session and type `dynwrap::test_docker_installation(detailed = TRUE)`:
```{r}
dynwrap::test_docker_installation(detailed = TRUE)
```
If it is `TRUE` in the result, that means Docker is successfully installed.

### Download simpipe Docker Image

Use `babelwhale::pull_container("duohongrui/simpipe")` command to pull simpipe Docker image in R (it will take 3-10 mins):
```{r eval=FALSE}
# install.packages("babelwhale")
babelwhale::pull_container("duohongrui/simpipe")
```

Once the image has been downloaded, check the image list:
```{r}
babelwhale::list_docker_images("duohongrui/simpipe")
```


### Install simpipe2docker

Get the simpipe2docker package from github by:
```{r eval=FALSE}
devtools::install_github("duohongrui/simpipe2docker")
```

Load the package and check whether it is installed successfully:
```{r eval=FALSE}
library(simpipe2docker)
```
Until now, you have successfully established the programming environment 👍👍👍. Go to the [ next page](/programming/2-estimation).👈️